---
title: "Baltimore City Employees' Salaries"
output: 
html_notebook: default
---

1. Load the necessary package(s)
```{r}

```

2. Load the data into a variable called "employees"
```{r}

```


## INTEGRITY CHECKS

Is any individual in here twice? Can you figure out why?
```{r}


```

How many years of data do we have?
```{r}

```

What are the min and max annual salary given out in 2020? 
```{r}

```

Which jobs get paid $0?
```{r}

```

How clean are the job titles? Are there a lot of duplicates? 
```{r}


```

Clean up the JobTitles by making everything lowercase
(hint: use the mutate() function to overwrite the current JobTitle field, using the function str_to_lower())
```{r}

```

Take a look at agency names; how clean are they? 
```{r}

```

## QUESTIONS OF THE DATA

Who's the highest paid employee in FY2020?
```{r}

```


Which agency has the highest average salary in 2020? 
```{r}

```

How many people work in the police department in 2020? 
```{r}

```

How many are "police officers"?
```{r}


```

What is the police officers' total salary? Average salary?
```{r}


```
