---
title: "R Basics"
output: html_notebook
---

This is a chance to practice writing some of the code we've covered so far. 
You can reference our "R Dictionary" here:
[https://github.com/eklucas/ADJS2022/blob/main/tipsheets/R-dictionary.md](https://github.com/eklucas/ADJS2022/blob/main/tipsheets/R-dictionary.md)


Create a *character vector* that lists your five favorite foods (or movies, TV shows, books, albums, whatever) and then print that vector. 
Remember that you have to give it a name.
```{r}

```

Create a *numeric vector* with every whole number between 10 and 20. 
```{r}

```

Pull out the seventh item in that vector.
```{r}

```

Create a data frame with two columns: one for the first names of everyone in this class and one for the last names.
```{r}

```

Pull out the column of last names.
```{r}

```

Pull out the second item in the last names column.
```{r}

```

Import the tidyverse package
```{r}

```

Import the MLB 2021 salaries data file
```{r}

```

Show which players play 1B (first base)
```{r}

```

Calculate their average salary
```{r}

```

Calculate the average salary per position; where do first basemen rank?
```{r}

```


