---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```

Downloaded from https://github.com/utdata/rwd-billboard-data on April 4, 2022.
Billboard Hot 100 data - top 100 songs every week. 

```{r}
billboard <- read_csv("data/hot100_kaggle_195808_20211106.txt")

billboard <- billboard %>% clean_names()
```


Rows: 330,087
What is one row: one song x one week
Date range? 8/4/1858 to 11/6/2021
```{r}
billboard %>% count(date) %>% arrange(desc(date))

billboard %>% summarise(range(date))
```

```{r}
billboard %>% filter(song=="Bohemian Rhapsody" & artist=="Queen")
```

```{r}
billboard %>% filter(song=="Yesterday" & grepl("Beatles", artist)) %>% arrange(peak_rank)

billboard %>% filter(song=="Yesterday" & grepl("Beatles", artist)) %>% summarise(min_rank = min(peak_rank))
```


Which song spent the longest time on Billboard Top 100?
Blinding Lights by The Weeknd
```{r}
billboard %>% arrange(desc(weeks_on_board))

billboard %>% group_by(artist,song) %>% 
  summarise(max_weeks = max(weeks_on_board)) %>% 
  arrange(desc(max_weeks))
```

Which artist has the most songs that hit the Top 100? 
Glee Cast with 183 songs. 
What are they?
```{r}
billboard %>% group_by(artist) %>% 
  summarise(num_songs = n_distinct(song)) %>%
  arrange(desc(num_songs))

billboard %>% filter(artist=="Glee Cast") %>%
  group_by(song) %>%
  summarise(rank = min(peak_rank)) %>%
  arrange(rank)
```

Which artist has the most cumulative weeks (regardless of song)? 
Taylor Swift
```{r}
billboard %>% count(artist) %>% arrange(desc(n))
```

How many songs has Beyonce had on the Top 100? 
```{r}
billboard %>% filter(grepl("Beyonce", artist) & !grepl("Featuring Beyonce", artist) & peak_rank==1) %>%
  distinct(song)
```



