---
title: "Campaign Finance - Feb 28"
output: 
html_notebook: default
---
#Introduction: 
Each candidate for federal office (U.S. House of Representatives, U.S. Senate, President) must register with the Federal Election Commission (FEC) and form a Political Action Committee (PAC) to receive contributions. Candidates can have several different kinds of PACs, but everyone has a "principal campaign committee" through which they receive money.

It's an election year: every Representative is up for re-election in an even year, since Reps serve 2-year terms. One third of the Senate is up for re-election every two years because Sens serve 6-year terms. Candidates can receive money even if they're not running for re-election, but individuals are limited in how much they can give a candidate each 2-year cycle: $2900.

Missouri has 8 House seats and 2 Senate seats; only 1 Senate seat is up for re-election in 2022; Roy Blunt is not running for re-election so it's an open seat. Josh Hawley won't be up for re-election until 2024.

```{r}
library(tidyverse)
```

Load the three tables: contributions, candidates, committees
```{r}
contributions <- read_csv("data/campfin/mo_contributions.csv")
candidates <- read_csv("data/campfin/candidates.csv")
committees <- read_csv("data/campfin/committees.csv")
```

#Documentation for these three tables: 

**Individual contributions**:
https://www.fec.gov/campaign-finance-data/contributions-individuals-file-description/
**Candidates**: 
https://www.fec.gov/campaign-finance-data/candidate-master-file-description/
**Committees**: 
https://www.fec.gov/campaign-finance-data/committee-master-file-description/


# INTEGRITY CHECKS
Contributions data: 
```{r}
contributions %>% 
  count(cmte_id) %>%
  inner_join(committees, by="cmte_id") %>%
  arrange(desc(n))
```

```{r}
contributions %>% group_by(transaction_tp) %>%
  summarise(total = sum(transaction_amt)) %>%
  arrange(desc(total))
```


# QUESTIONS OF THE DATA

Professors from MU who donated: 
```{r}
contributions %>% filter((grepl("UNIV.*MISSOURI", employer) | 
                           grepl("UNIV.*MO", employer)) & 
                           employer!="UNIV OF CENTRAL MO") %>%
  summarise(total = sum(transaction_amt))
```

```{r}
candidates %>% filter(election_yr=="2022") %>%
  count(party) %>%
  arrange(desc(n))
```

```{r}
candidates %>% 
  filter(election_yr=="2022" & office_st=="MO" & office=="S") %>%
  inner_join(contributions, by=c("pcc"="cmte_id")) %>%
  group_by(party) %>%
  summarise(total = sum(transaction_amt)) %>%
  arrange(desc(total))
```

```{r}
contributions %>% count(occupation) %>% filter(grepl("CEO", occupation))
```

