---
title: "NYC Crashes"
output:
html_notebook: default

---
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(RSocrata)   # you'll need to run install.packages("RSocrata") first
```


We'll be directly downloading some data from the NYC Open Data Portal
Here's the page: [https://data.cityofnewyork.us/Public-Safety/Motor-Vehicle-Collisions-Crashes/h9gi-nx95](https://data.cityofnewyork.us/Public-Safety/Motor-Vehicle-Collisions-Crashes/h9gi-nx95)
* each row is a single crash event
* there are blanks in the "borough" column

Here's how to get the data:
* Click API
* Change the API Endpoint from JSON to CSV
* Copy

In order to make things a little easier on ourselves, we're going to narrow our request (the main file is quite large). If you read the API docs and try some things you can find ways to do this.



```{r}
data <- read.socrata("https://data.cityofnewyork.us/resource/h9gi-nx95.csv?borough=MANHATTAN")
```

Notes about the data: 
* 292,126 rows (out of 1.83 million) - 16%
* there are some blanks in the on_street_name, off_street_name, cross_street_name

Now, some questions!
