---
title: "UCD Deaths: Feb 7 - 16"
output: 
html_notebook: default
---

```{r}
library(tidyverse)
library(janitor)
```

###Download and import the data
We downloaded "Underlying Cause of Death" data by state, year and cause for 2018-2020 (the most recent year available) from [CDC Wonder](wonder.cdc.gov).

```{r}
deaths <- read_tsv("data/ucd-deaths-by-state-year-cause.txt")
```

###Things to deal with: 
Suppressed numbers in deaths, population
Deaths and population are character columns
Random hashtags in Cause List? (these are "rankable deaths" according to the Notes)
Remove notes from the bottom
Clean column names


Clean names:
```{r}
deaths <- deaths %>% clean_names()
```

Separate out the notes, but save them because they have valuable information:
```{r}
notes <- deaths %>% filter(!is.na(notes)) %>% select(notes)
deaths <- deaths %>% filter(is.na(notes))
```

Add new columns for deaths and population that are converted to numbers (NAs will be introduced where the original values are "Suppressed"):
```{r}
deaths <- deaths %>% mutate(num_deaths = as.numeric(deaths))
deaths <- deaths %>% mutate(num_pop = as.numeric(population))
```

Save both the notes and the cleaned up deaths data:
```{r}
write_csv(notes, "data/ucd-notes.csv")
write_csv(deaths, "data/ucd-clean.csv")
```

###Questions we asked in class: 
How many people die in the US every year? 
What caused the most deaths in Missouri? 
Which state reported the most deaths? 
How many people died from firearms? Drugs? Suicide? 
What is the most common cause of death for each year? 
How do Missouri common causes of death compare to national causes?
Compare common causes of death by state? 
Which causes of death is increasing from year to year? 
Compare infections diseases in Missouri to other states.

1. How many people die in the US every year? 
```{r}
deaths %>% 
  group_by(year) %>% 
  summarise(total_deaths = sum(num_deaths, na.rm=T))
```
2. What caused the most deaths in Missouri? 
```{r}
deaths %>% filter(state=="Missouri") %>% 
  group_by(icd_10_113_cause_list) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  arrange(desc(total_deaths))
```

What cause the most deaths in Missouri in 2020? 
```{r}
MO2020 <- deaths %>% filter(state=="Missouri" & year==2020) %>%
  select(icd_10_113_cause_list, num_deaths) %>%
  arrange(desc(num_deaths))
```

3. Which state reported the most deaths? 
```{r}
deaths %>% 
  group_by(state) %>% 
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  arrange(desc(total_deaths))
```
Which state had the highest death rate in 2020? 
```{r}
deaths %>%
  filter(year==2020) %>%
  group_by(state, num_pop) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  mutate(rate = total_deaths/num_pop*100000) %>%
  arrange(desc(rate))
```

4. How many people died from firearms? Drugs? Suicide? 
Firearms: 122,205
Suicide: 283,650
There is some overlap as some suicide cases are assisted by firearms
```{r}

deaths %>% count(icd_10_113_cause_list)%>% 
  filter(grepl("suicide|firearm|drug", icd_10_113_cause_list, ignore.case=TRUE)) # overdose deaths are not included here

# deaths by firearms:
deaths %>% filter(grepl("firearm", icd_10_113_cause_list, ignore.case=TRUE)) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T))

# deaths by firearms by year:
deaths %>% filter(grepl("firearm", icd_10_113_cause_list, ignore.case=TRUE)) %>%
  group_by(year) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) 

# deaths by suicide:
deaths %>% filter(grepl("suicide", icd_10_113_cause_list, ignore.case=TRUE)) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T))

```
Look for firearm and suicide together:
333,208
```{r}
deaths %>% 
  filter(grepl("suicide|firearm", icd_10_113_cause_list, ignore.case=TRUE)) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T))
```

5. What is the most common cause of death for each year?
Each year, major cardiovascular diseases are #1, followed by Diseases of heart and Malignant neoplasms.
```{r}
deaths %>% 
  filter(year=="2020") %>%
  group_by(icd_10_113_cause_list) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  arrange(desc(total_deaths))

deaths %>% 
  filter(year=="2019") %>%
  group_by(icd_10_113_cause_list) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  arrange(desc(total_deaths))

deaths %>% 
  filter(year=="2018") %>%
  group_by(icd_10_113_cause_list) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  arrange(desc(total_deaths))
```

6. How do Missouri common causes of death compare to national causes?
The top five causes are the same.
```{r}
national <- deaths %>% 
  group_by(icd_10_113_cause_list) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  arrange(desc(total_deaths))

Missouri <- deaths %>% 
  filter(state=="Missouri") %>%
  group_by(icd_10_113_cause_list) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  arrange(desc(total_deaths))
```


7. Compare common causes of death by state.
Because this question is fairly vague, it's hard to answer directly with just a few lines of code. It's easier to narrow the scope a bit, so let's compare rates for the top three causes by state for 2020 to see if something pops out: 
```{r}
deaths %>%
  filter(icd_10_113_cause_list %in% c("Major cardiovascular diseases (I00-I78)", 
                                      "#Diseases of heart (I00-I09,I11,I13,I20-I51)", 
                                      "#Malignant neoplasms (C00-C97)") 
         & year == "2020" & !is.na(num_deaths)) %>%
  mutate(rate = num_deaths/num_pop*100000) %>%
  select(state, icd_10_113_cause_list, rate) %>%
  pivot_wider(names_from = icd_10_113_cause_list, values_from=rate)
```


8. Which causes of death are increasing from year to year? 
```{r}
chg_in_cause <- deaths %>%
  group_by(icd_10_113_cause_list, year) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  pivot_wider(names_from=year, values_from=total_deaths) %>%
  mutate(pct_chg = round((`2020`-`2018`)/`2018`, digits=4)) %>%
  arrange(desc(pct_chg))
```


Create a population table (we'll need this for the next question):
```{r}
pop <- deaths %>% distinct(state, year, num_pop) %>% filter(!is.na(num_pop))
```


9. Compare infections diseases in Missouri to other states.
First off, it's not easy to identify infectious diseases in this data: HIV is an infectious disease, but not labeled as such. In order to truly identify all causes that are considered infectious diseases, we'd have to consult an expert. For our purposes, we'll consider a few basics: HIV, respiratory infections, and other unspecified infectious and parasitic diseases.
It's probably best to calculate rates for infections diseases by state for fair comparison.

Missouri is 25th (right in the middle) in 2018 but 18th in 2020. 
```{r}
# test the keywords to search for:
deaths %>% 
  count(icd_10_113_cause_list) %>%
  filter(grepl("HIV|respiratory infection|parasitic disease", icd_10_113_cause_list))

# calculate rates by state:
deaths %>% 
  filter(grepl("HIV|respiratory infection|parasitic disease", icd_10_113_cause_list)) %>%
  group_by(state, year) %>%
  summarise(total_deaths = sum(num_deaths, na.rm=T)) %>%
  inner_join(pop, by=c("state","year")) %>%
  mutate(rate = total_deaths/num_pop*100000) %>%
  arrange(year, desc(rate))
```




